---
title: "Hypothesis Testing"
format: html
editor: visual
---

## CPS Library Pull

```{r}
library(tidyverse)
```

```{r}
library(probstats4econ)
cps%>%filter(statefips%in%c("VT","CA"))%>%glimpse()

```

```{r}
vt_ca <- cps %>%
  filter(statefips %in% c("VT", "CA"))
```

## One Sided T-test

```{r}
t.test(age ~ statefips,
       data = vt_ca,
       alternative = "greater")
```

::: textbox
**t value:** indicates that the mean of data set "ca" (California) is lower than that of data set "vt" (Vermont).

**p-value:** the value is high indicating that there is a high chance the null hypothesis is true.

**alternative hypothesis:** the test was run to see if the result of ca-vt was less than or equal to 0, and a negative difference would indicate that the mean of ca is less than vt.
:::

## Box and Whisker Plot

```{r}
vt <- vt_ca$age[vt_ca$statefips == "VT"]

boxplot(vt,
        main = "Vermont Age Distribution",
        ylab = "Age",
        xlab = "")  # empty x-axis label


```

```{r}
ca <- vt_ca$age[vt_ca$statefips == "CA"]

boxplot(ca,
        main = "California Age Distribution",
        ylab = "Age",
        xlab = "")

```

## T-Distribution Plot w/ Critical Value

```{r}
# Your t-test values
t_stat <- -0.44845
df <- 40.22

# Plot t-distribution
curve(dt(x, df = df), from = -4, to = 4, ylab = "Density")

# 95% quantile for one-tailed test (right tail)
abline(v = qt(0.95, df = df), col = "red", lty = 2)

# t-statistic
abline(v = t_stat, col = "blue", lwd = 2)

```
